---
title: "Political Lean Classification with Naive Bayes"
author: "Chica Morrow and Bryce Samwel"
date: "11/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Many people and tables discuss the relationship between political leaning and a given demographic. However, what we are attempting to do is predict someone's political leaning based on many demographic features. Presently, we are considering a person's sex, race, generation, community type, income, education, and marital status. To accomplish this we are using a Naive Bayes Classifier and survey data from the Pew Research Center. Before we jump into the code, we should discuss the Naive Bayes Classifier for those that aren't familiar.
\vspace{12 pt}

## Naive Bayes

The Naive Bayes Classifier is a probabilistic classifier based on the application of Bayes' theorem. One of the major assumptions made by this classifier is that the various features are independent. In almost all cases that won't be true, but it's worth noting before we continue. Considering that Bayes' theorem is central to this classifier, let's take a look at the mathematical formulation and make some sense of it.

$$
p(A|B) = \frac{p(A)p(B|A)}{p(B)}
$$
For example, if A is "lean republican" and B is "is married", Bayes' theorem says that the probability a person will lean republican given that they are married is equal to the probability that a person will lean republican, multiplied by the probability that a person will be married given that they lean republican, divided by the probability that a person will be married. Putting this equation into Bayesian probability terminology, it can be read as follows:

$$
\text{posterior}=\frac{\text{prior}*\text{likelihood}}{\text{evidence}}
$$
We refer to the prior in this instance as the class prior, where the classes are the political leanings, and the evidence as the predictor prior, where the predictors are the demographic features of a given individual.

Okay, so we now know how to predict the probability of someone leaning a certain way given a single demographic. As it turns out, the simple classifier doesn't get much more complicated than this. Given many features, we can calculate a class "score" by multiplying the likelihoods for that class as well as the prior for that class. The class with the highest "score" is selected as the prediction. You can see an example of how this is applied with our code at the end of this paper.
\vspace{12 pt}

## A Brief Discussion of the Data

Normally, we would use a dataset of individuals with information on the features and political leaning of each of them. However, we couldn't find such a dataset very easily. Instead, we found a table from the Pew Research Center. It's titled "2016 Party Identification Detailed Tables." Oddly enough, the table gives us the posteriors and class priors. Additionally, one should only build a model based on a single dataset. However, the predictor priors from the Pew Research Center table differ a little bit from census data. We instead use census data from 2016 for our predictor priors. Ultimately, we need to calculate the likelihoods manually using Bayes' theorem. 

## Calculating the Likelihoods

This section is very code-heavy and is primarily the hand-encoding of each of the priors and the posteriors. First, let's look at the posteriors.
\vspace{12 pt}

```{r posteriors}
#Probabilities given sex

probRepGivenMale <- 0.51
probDemGivenMale <- 0.41
probIndGivenMale <- 0.08

probRepGivenFemale <- 0.38
probDemGivenFemale <- 0.54
probIndGivenFemale <- 0.08

#Probabilities given race/ethnicity

probRepGivenWhite <- 0.54
probDemGivenWhite <- 0.39
probIndGivenWhite <- 0.07

probRepGivenBlack <- 0.07
probDemGivenBlack <- 0.87
probIndGivenBlack <- 0.06

probRepGivenHispanic <- 0.27
probDemGivenHispanic <- 0.63
probIndGivenHispanic <- 0.10

probRepGivenAsian <- 0.27
probDemGivenAsian <- 0.66
probIndGivenAsian <- 0.07

#Probabilities given generation

probRepGivenMillenial <- 0.36
probDemGivenMillenial <- 0.57
probIndGivenMillenial <- 0.07

probRepGivenGenX <- 0.42
probDemGivenGenX <- 0.48
probIndGivenGenX <- 0.09

probRepGivenBoomer <- 0.49
probDemGivenBoomer <- 0.45
probIndGivenBoomer <- 0.06

probRepGivenSilent <- 0.53
probDemGivenSilent <- 0.40
probIndGivenSilent <- 0.07

#Probabilities given community type

probRepGivenUrban <- 0.33
probDemGivenUrban <- 0.60
probIndGivenUrban <- 0.07

probRepGivenSuburban <- 0.48
probDemGivenSuburban <- 0.44
probIndGivenSuburban <- 0.07

probRepGivenRural <- 0.55
probDemGivenRural <- 0.37
probIndGivenRural <- 0.08

#Probabilities given income

probRepGiven75plus <- 0.49
probDemGiven75plus <- 0.45
probIndGiven75plus <- 0.06

probRepGiven30to75 <- 0.48
probDemGiven30to75 <- 0.45
probIndGiven30to75 <- 0.07

probRepGivenlessthan30 <- 0.32
probDemGivenlessthan30 <- 0.60
probIndGivenlessthan30 <- 0.08

#Probabilities given education

probRepGivenPostgrad <- 0.35
probDemGivenPostgrad <- 0.60
probIndGivenPostgrad <- 0.05

probRepGivenCollegeGrad <- 0.44
probDemGivenCollegeGrad <- 0.50
probIndGivenCollegeGrad <- 0.06

probRepGivenSomeCollege <- 0.46
probDemGivenSomeCollege <- 0.45
probIndGivenSomeCollege <- 0.09

probRepGivenHighSchool <- 0.45
probDemGivenHighSchool <- 0.46
probIndGivenHighSchool <- 0.08

#Probabilities given marital status

probRepGivenMarried <- 0.51
probDemGivenMarried <- 0.44
probIndGivenMarried <- 0.06

probRepGivenUnmarried <- 0.36
probDemGivenUnmarried <- 0.56
probIndGivenUnmarried <- 0.08

```

Now, let's see the class priors.
\vspace{12 pt}

```{r class_priors}
#Class priors

probRep <- 0.44
probDem <- 0.48
probInd <- 0.08

```

And the predictor priors...
\vspace{12 pt}

```{r predictor_priors}
#Sex priors

probMale <- 0.492
probFemale <- 0.508

#Race/ethnicity priors

probWhite <- 0.62
probBlack <- 0.123
probHispanic <- 0.173
probAsian <- 0.052

#Generation Priors

probMillenial <- 0.197
probGenX <- 0.23
probBoomer <- 0.408
probSilent <- 0.145

#Community Type priors
probUrban <- 0.336
probSuburban <- 0.466
probRural <- 0.171
probHispanic <- 0.125

#Income priors

prob75plus <- 0.452
prob30to75 <- 0.350
problessthan30 <- 0.198

#Education priors

probPostgrad <- 0.159
probCollegeGrad <- 0.329
probSomeCollege <- 0.278
probHighSchool <- 0.229

#Marital Status priors

probMarried <- 0.574
probUnmarried <- 0.426

```

Now that we have all of these, we can use Bayes' theorem to calculate each of the likelihoods which will be critical in making our predictions.
\vspace{12 pt}

```{r likelihoods}
##Likelihoods given Rep

#by sex
probMaleGivenRep <- probRepGivenMale*probMale/probRep
probFemaleGivenRep <- probRepGivenFemale*probFemale/probRep

#by race
probWhiteGivenRep <- probRepGivenWhite*probWhite/probRep
probBlackGivenRep <- probRepGivenBlack*probBlack/probRep
probHispanicGivenRep <- probRepGivenHispanic*probHispanic/probRep
probAsianGivenRep <- probRepGivenAsian*probAsian/probRep

#by generation
probMillenialGivenRep <- probRepGivenMillenial*probMillenial/probRep
probGenXGivenRep <- probRepGivenGenX*probGenX/probRep
probBoomerGivenRep <- probRepGivenBoomer*probBoomer/probRep
probSilentGivenRep <- probRepGivenSilent*probSilent/probRep

#by community type
probUrbanGivenRep <- probRepGivenUrban*probUrban/probRep
probSuburbanGivenRep <- probRepGivenSuburban*probSuburban/probRep
probRuralGivenRep <- probRepGivenRural*probRural/probRep

#by income
prob75plusGivenRep <- probRepGiven75plus*prob75plus/probRep
prob30to75GivenRep <- probRepGiven30to75*prob30to75/probRep
problessthan30GivenRep <- probRepGivenlessthan30*problessthan30/probRep

#by education
probPostgradGivenRep <- probRepGivenPostgrad*probPostgrad/probRep
probCollegeGradGivenRep <- probRepGivenCollegeGrad*probCollegeGrad/probRep
probSomeCollegeGivenRep <- probRepGivenSomeCollege*probSomeCollege/probRep
probHighSchoolGivenRep <- probRepGivenHighSchool*probHighSchool/probRep

#by marital status
probMarriedGivenRep <- probRepGivenMarried*probMarried/probRep
probUnmarriedGivenRep <- probRepGivenUnmarried*probUnmarried/probRep

##Likelihoods given Dem

#by sex
probMaleGivenDem <- probDemGivenMale*probMale/probDem
probFemaleGivenDem <- probDemGivenFemale*probFemale/probDem

#by race
probWhiteGivenDem <- probDemGivenWhite*probWhite/probDem
probBlackGivenDem <- probDemGivenBlack*probBlack/probDem
probHispanicGivenDem <- probDemGivenHispanic*probHispanic/probDem
probAsianGivenDem <- probDemGivenAsian*probAsian/probDem

#by generation
probMillenialGivenDem <- probDemGivenMillenial*probMillenial/probDem
probGenXGivenDem <- probDemGivenGenX*probGenX/probDem
probBoomerGivenDem <- probDemGivenBoomer*probBoomer/probDem
probSilentGivenDem <- probDemGivenSilent*probSilent/probDem

#by community type
probUrbanGivenDem <- probDemGivenUrban*probUrban/probDem
probSuburbanGivenDem <- probDemGivenSuburban*probSuburban/probDem
probRuralGivenDem <- probDemGivenRural*probRural/probDem

#by income
prob75plusGivenDem <- probDemGiven75plus*prob75plus/probDem
prob30to75GivenDem <- probDemGiven30to75*prob30to75/probDem
problessthan30GivenDem <- probDemGivenlessthan30*problessthan30/probDem

#by education
probPostgradGivenDem <- probDemGivenPostgrad*probPostgrad/probDem
probCollegeGradGivenDem <- probDemGivenCollegeGrad*probCollegeGrad/probDem
probSomeCollegeGivenDem <- probDemGivenSomeCollege*probSomeCollege/probDem
probHighSchoolGivenDem <- probDemGivenHighSchool*probHighSchool/probDem

#by marital status
probMarriedGivenDem <- probDemGivenMarried*probMarried/probDem
probUnmarriedGivenDem <- probDemGivenUnmarried*probUnmarried/probDem

##Likelihoods given Ind

#by sex
probMaleGivenInd <- probIndGivenMale*probMale/probInd
probFemaleGivenInd <- probIndGivenFemale*probFemale/probInd

#by race
probWhiteGivenInd <- probIndGivenWhite*probWhite/probInd
probBlackGivenInd <- probIndGivenBlack*probBlack/probInd
probHispanicGivenInd <- probIndGivenHispanic*probHispanic/probInd
probAsianGivenInd <- probIndGivenAsian*probAsian/probInd

#by generation
probMillenialGivenInd <- probIndGivenMillenial*probMillenial/probInd
probGenXGivenInd <- probIndGivenGenX*probGenX/probInd
probBoomerGivenInd <- probIndGivenBoomer*probBoomer/probInd
probSilentGivenInd <- probIndGivenSilent*probSilent/probInd

#by community type
probUrbanGivenInd <- probIndGivenUrban*probUrban/probInd
probSuburbanGivenInd <- probIndGivenSuburban*probSuburban/probInd
probRuralGivenInd <- probIndGivenRural*probRural/probInd

#by income
prob75plusGivenDem <- probDemGiven75plus*prob75plus/probDem
prob30to75GivenDem <- probDemGiven30to75*prob30to75/probDem
problessthan30GivenDem <- probDemGivenlessthan30*problessthan30/probDem

#by education
probPostgradGivenInd <- probIndGivenPostgrad*probPostgrad/probInd
probCollegeGradGivenInd <- probIndGivenCollegeGrad*probCollegeGrad/probInd
probSomeCollegeGivenInd <- probIndGivenSomeCollege*probSomeCollege/probInd
probHighSchoolGivenInd <- probIndGivenHighSchool*probHighSchool/probInd

#by marital status
probMarriedGivenInd <- probIndGivenMarried*probMarried/probInd
probUnmarriedGivenInd <- probIndGivenUnmarried*probUnmarried/probInd

```

## An Example Prediction

If we come up with an example person, we can see which class (political leaning) they are most likely to fall into! Let's suppose that we are interested in knowing a black woman's leaning. It's as simple as multiplying the prior for a given class by the likelihoods for that person given that class. The class with the highest outcome is our prediction! Let's do the calculations.
\vspace{12 pt}

```{r Black_Woman_Prediction}
#Republican

probRep*probBlackGivenRep*probFemaleGivenRep

#Democrat

probDem*probBlackGivenDem*probFemaleGivenDem

#Independent

probInd*probBlackGivenInd*probFemaleGivenInd
```

As we can see, the value for leaning democratic is higher than all of the other values, so we would predict that a black woman leans democratic based on this simple model!
\vspace{12 pt}