---
title: "Political Lean Classification"
author: "Chica Morrow and Bryce Samwel"
date: "11/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We're gonna try to tell people whether they lean democratic or republican with a
Naive Bayes classifier. We will start with a very simple model, only looking at 
sex and race.

First, let's look at the posteriors from the Pew Research Center.
\vspace{12 pt}

```{r posteriors}
#Probabilities given sex

probRepGivenMale <- 0.51
probDemGivenMale <- 0.41
probIndGivenMale <- 0.08

probRepGivenFemale <- 0.38
probDemGivenFemale <- 0.54
probIndGivenFemale <- 0.08

#Probabilities given race/ethnicity

probRepGivenWhite <- 0.54
probDemGivenWhite <- 0.39
probIndGivenWhite <- 0.07

probRepGivenBlack <- 0.07
probDemGivenBlack <- 0.87
probIndGivenBlack <- 0.06

probRepGivenHispanic <- 0.27
probDemGivenHispanic <- 0.63
probIndGivenHispanic <- 0.10

#Probabilities given generation

probRepGivenMillennial <- .36
probDemGivenMillennial <- .57
probIndGivenMillennial <- .41

probRepGivenGenX <- .42
probDemGivenGenX <- .48
probDemGivenGenX <- .36

probRepGivenBabyBoomer <- .49
probDemGivenBabyBoomer <- .45
probDemGivenBabyBoomer <- .30

probRepGivenSilent <- .53
probDemGivenSilent <- .40
probIndGivenSilent <- .23

#Probabilities given income

probRepGiven75plus <- 49
probDemGiven75plus <- 45
probIndGiven75plus <- 34

probRepGiven30to75 <- 48
probDemGiven30to75 <- 45
probIndGiven30to75 <- 34

probRepGivenlessthan30 <- 32
probDemGivenlessthan30 <- 60
probIndGivenlessthan30 <- 34

```

Now, let's see the class priors.
\vspace{12 pt}

```{r class_priors}
#Class priors

probRep <- 0.44
probDem <- 0.48
probInd <- 0.08

```

And the predictor priors...
\vspace{12 pt}

```{r predictor_priors}
#Sex priors

probMale <- 0.492
probFemale <- 0.508

#Race/ethnicity priors

probWhite <- 0.62
probBlack <- 0.123
probHispanic <- 0.125

#Generation priors

probMillennial <- 0.201
probGenX <- 0.235
probBabyBoomer <- 0.417
probSilent <- 0.148

#Income priors

prob75plus <- 0.452
prob30to75 <- 0.350
problessthan30 <- 0.198



```

Now that we have all of these, we can used the naive bayes theorem to calculate 
each of the likelihoods which will be critical in making our predictions.
\vspace{12 pt}

```{r likelihoods}
##Likelihoods given Rep

#by sex
probMaleGivenRep <- probRepGivenMale*probMale/probRep
probFemaleGivenRep <- probRepGivenFemale*probFemale/probRep

#by race
probWhiteGivenRep <- probRepGivenWhite*probWhite/probRep
probBlackGivenRep <- probRepGivenBlack*probBlack/probRep
probHispanicGivenRep <- probRepGivenHispanic*probHispanic/probRep

#by generation
probMillennialGivenRep <- probRepGivenMillennial/probRep
probGenXGivenRep <- probRepGivenGenX/probRep
probBabyBoomerGivenRep <- probRepGivenBabyBoomer/probRep
probSilentGivenRep <- probRepGivenSilent/probRep

#by income
prob75plusGivenRep <- probRepGiven75plus/probRep
prob30to75GivenRep <- probRepGiven30to75/probRep
problessthan30GivenRep <- probRepGivenlessthan30/probRep

##Likelihoods given Dem

#by sex
probMaleGivenDem <- probDemGivenMale*probMale/probDem
probFemaleGivenDem <- probDemGivenFemale*probFemale/probDem

#by race
probWhiteGivenDem <- probDemGivenWhite*probWhite/probDem
probBlackGivenDem <- probDemGivenBlack*probBlack/probDem
probHispanicGivenDem <- probDemGivenHispanic*probHispanic/probDem

#by generation
probMillennialGivenDem <- probDemGivenMillennial/probDem
probGenXGivenDem <- probDemGivenGenX/probDem
probBabyBoomerGivenDem <- probDemGivenBabyBoomer/probDem
probSilentGivenDem <- probDemGivenSilent/probDem

#by income
prob75plusGivenDem <- probDemGiven75plus/probDem
prob30to75GivenDem <- probDemGiven30to75/probDem
problessthan30GivenDem <- probDemGivenlessthan30/probDem

##Likelihoods given Ind

#by sex
probMaleGivenInd <- probIndGivenMale*probMale/probInd
probFemaleGivenInd <- probIndGivenFemale*probFemale/probInd

#by race
probWhiteGivenInd <- probIndGivenWhite*probWhite/probInd
probBlackGivenInd <- probIndGivenBlack*probBlack/probInd
probHispanicGivenInd <- probIndGivenHispanic*probHispanic/probInd

#by generation
probMillennialGivenDem <- probDemGivenMillennial/probDem
probGenXGivenDem <- probDemGivenGenX/probDem
probBabyBoomerGivenDem <- probDemGivenBabyBoomer/probDem
probSilentGivenDem <- probDemGivenSilent/probDem

#by income
prob75plusGivenDem <- probDemGiven75plus/probDem
prob30to75GivenDem <- probDemGiven30to75/probDem
problessthan30GivenDem <- probDemGivenlessthan30/probDem
```

So, if we come up with an example person, we can see which class (political 
leaning) they are most likely to fall into! Let's suppose that we are interested 
in knowing a black woman's leaning. It's as simple as multiplying the prior for 
a given class by the likelihoods for that person given that class. The class with 
the highest outcome is our prediction! Let's do the calculations.
\vspace{12 pt}

```{r Black_Woman_Prediction}
#Republican

probRep*probBlackGivenRep*probFemaleGivenRep

#Democrat

probDem*probBlackGivenDem*probFemaleGivenDem

#Independent

probInd*probBlackGivenInd*probFemaleGivenInd
```

As we can see, the value for leaning democratic is higher than all of the other 
values, so we would predict that a black woman leans democratic based on this 
simple model! Now that might have been rather obvious, but what if we are 
interested in how a white woman might lean...
\vspace{12 pt}

```{r White_Woman_Prediction}
#Republican

probRep*probWhiteGivenRep*probFemaleGivenRep

#Democrat

probDem*probWhiteGivenDem*probFemaleGivenDem

#Independent

probInd*probWhiteGivenInd*probFemaleGivenInd
```