---
title: "Political Lean Classification"
author: "Chica Morrow and Bryce Samwel"
date: "11/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We're gonna try to tell people whether they lean democratic or republican with a
Naive Bayes classifier. We will start with a very simple model, only looking at 
sex and race.

First, let's look at the posteriors from the Pew Research Center.
\vspace{12 pt}

```{r posteriors}
#Probabilities given sex

probRepGivenMale <- 0.51
probDemGivenMale <- 0.41
probIndGivenMale <- 0.08

probRepGivenFemale <- 0.38
probDemGivenFemale <- 0.54
probIndGivenFemale <- 0.08

#Probabilities given race/ethnicity

probRepGivenWhite <- 0.54
probDemGivenWhite <- 0.39
probIndGivenWhite <- 0.07

probRepGivenBlack <- 0.07
probDemGivenBlack <- 0.87
probIndGivenBlack <- 0.06

probRepGivenHispanic <- 0.27
probDemGivenHispanic <- 0.63
probIndGivenHispanic <- 0.10
```

Now, let's see the class priors.
\vspace{12 pt}

```{r class_priors}
#Class priors

probRep <- 0.44
probDem <- 0.48
probInd <- 0.08

```

And the predictor priors...
\vspace{12 pt}

```{r predictor_priors}
#Sex priors

probMale <- 0.492
probFemale <- 0.508

#Race/ethnicity priors

probWhite <- 0.62
probBlack <- 0.123
probHispanic <- 0.125
```

Now that we have all of these, we can used the naive bayes theorem to calculate 
each of the likelihoods which will be critical in making our predictions.
\vspace{12 pt}

```{r likelihoods}
#Likelihoods given Rep

probMaleGivenRep <- probRepGivenMale*probMale/probRep
probFemaleGivenRep <- probRepGivenFemale*probFemale/probRep

probWhiteGivenRep <- probRepGivenWhite*probWhite/probRep
probBlackGivenRep <- probRepGivenBlack*probBlack/probRep
probHispanicGivenRep <- probRepGivenHispanic*probHispanic/probRep

#Likelihoods given Dem

probMaleGivenDem <- probDemGivenMale*probMale/probDem
probFemaleGivenDem <- probDemGivenFemale*probFemale/probDem

probWhiteGivenDem <- probDemGivenWhite*probWhite/probDem
probBlackGivenDem <- probDemGivenBlack*probBlack/probDem
probHispanicGivenDem <- probDemGivenHispanic*probHispanic/probDem

#Likelihoods given Ind

probMaleGivenInd <- probIndGivenMale*probMale/probInd
probFemaleGivenInd <- probIndGivenFemale*probFemale/probInd

probWhiteGivenInd <- probIndGivenWhite*probWhite/probInd
probBlackGivenInd <- probIndGivenBlack*probBlack/probInd
probHispanicGivenInd <- probIndGivenHispanic*probHispanic/probInd
```

So, if we come up with an example person, we can see which class (political 
leaning) they are most likely to fall into! Let's suppose that we are interested 
in knowing a black woman's leaning. It's as simple as multiplying the prior for 
a given class by the likelihoods for that person given that class. The class with 
the highest outcome is our prediction! Let's do the calculations.
\vspace{12 pt}

```{r Black_Woman_Prediction}
#Republican

probRep*probBlackGivenRep*probFemaleGivenRep

#Democrat

probDem*probBlackGivenDem*probFemaleGivenDem

#Independent

probInd*probBlackGivenInd*probFemaleGivenInd
```

As we can see, the value for leaning democratic is higher than all of the other 
values, so we would predict that a black woman leans democratic based on this 
simple model! Now that might have been rather obvious, but what if we are 
interested in how a white woman might lean...
\vspace{12 pt}

```{r White_Woman_Prediction}
#Republican

probRep*probWhiteGivenRep*probFemaleGivenRep

#Democrat

probDem*probWhiteGivenDem*probFemaleGivenDem

#Independent

probInd*probWhiteGivenInd*probFemaleGivenInd
```