---
title: "Political Lean Classification"
author: "Chica Morrow and Bryce Samwel"
date: "11/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We're gonna try to tell people whether they lean democratic or republican with a Naive Bayes classifier. We will look at sex, race, generation, and community type.

First, let's look at the posteriors from the Pew Research Center.
\vspace{12 pt}

```{r posteriors}
#Probabilities given sex

probRepGivenMale <- 0.51
probDemGivenMale <- 0.41
probIndGivenMale <- 0.08

probRepGivenFemale <- 0.38
probDemGivenFemale <- 0.54
probIndGivenFemale <- 0.08

#Probabilities given race/ethnicity

probRepGivenWhite <- 0.54
probDemGivenWhite <- 0.39
probIndGivenWhite <- 0.07

probRepGivenBlack <- 0.07
probDemGivenBlack <- 0.87
probIndGivenBlack <- 0.06

probRepGivenHispanic <- 0.27
probDemGivenHispanic <- 0.63
probIndGivenHispanic <- 0.10

probRepGivenAsian <- 0.27
probDemGivenAsian <- 0.66
probIndGivenAsian <- 0.07

#Probabilities given generation

probRepGivenMillenial <- 0.36
probDemGivenMillenial <- 0.57
probIndGivenMillenial <- 0.07

probRepGivenGenX <- 0.42
probDemGivenGenX <- 0.48
probIndGivenGenX <- 0.09

probRepGivenBoomer <- 0.49
probDemGivenBoomer <- 0.45
probIndGivenBoomer <- 0.06

probRepGivenSilent <- 0.53
probDemGivenSilent <- 0.40
probIndGivenSilent <- 0.07

#Probabilities given community type

probRepGivenUrban <- 0.33
probDemGivenUrban <- 0.60
probIndGivenUrban <- 0.07

probRepGivenSuburban <- 0.48
probDemGivenSuburban <- 0.44
probIndGivenSuburban <- 0.07

probRepGivenRural <- 0.55
probDemGivenRural <- 0.37
probIndGivenRural <- 0.08

#Probabilities given income

probRepGiven75plus <- 0.49
probDemGiven75plus <- 0.45
probIndGiven75plus <- 0.06

probRepGiven30to75 <- 0.48
probDemGiven30to75 <- 0.45
probIndGiven30to75 <- 0.07

probRepGivenlessthan30 <- 0.32
probDemGivenlessthan30 <- 0.60
probIndGivenlessthan30 <- 0.08

#Probabilities given education

probRepGivenPostgrad <- 0.35
probDemGivenPostgrad <- 0.60
probIndGivenPostgrad <- 0.05

probRepGivenCollegeGrad <- 0.44
probDemGivenCollegeGrad <- 0.50
probIndGivenCollegeGrad <- 0.06

probRepGivenSomeCollege <- 0.46
probDemGivenSomeCollege <- 0.45
probIndGivenSomeCollege <- 0.09

probRepGivenHighSchool <- 0.45
probDemGivenHighSchool <- 0.46
probIndGivenHighSchool <- 0.08

#Probabilities given marital status

probRepGivenMarried <- 0.51
probDemGivenMarried <- 0.44
probIndGivenMarried <- 0.06

probRepGivenUnmarried <- 0.36
probDemGivenUnmarried <- 0.56
probIndGivenUnmarried <- 0.08

```

Now, let's see the class priors.
\vspace{12 pt}

```{r class_priors}
#Class priors

probRep <- 0.44
probDem <- 0.48
probInd <- 0.08

```

And the predictor priors...
\vspace{12 pt}

```{r predictor_priors}
#Sex priors

probMale <- 0.492
probFemale <- 0.508

#Race/ethnicity priors

probWhite <- 0.62
probBlack <- 0.123
probHispanic <- 0.173
probAsian <- 0.052

#Generation Priors

probMillenial <- 0.197
probGenX <- 0.23
probBoomer <- 0.408
probSilent <- 0.145

#Community Type priors
probUrban <- 0.336
probSuburban <- 0.466
probRural <- 0.171
probHispanic <- 0.125

#Income priors

prob75plus <- 0.452
prob30to75 <- 0.350
problessthan30 <- 0.198

#Education priors

probPostgrad <- 0.159
probCollegeGrad <- 0.329
probSomeCollege <- 0.278
probHighSchool <- 0.229

#Marital Status priors

probMarried <- 0.574
probUnmarried <- 0.426

```

Now that we have all of these, we can used the naive bayes theorem to calculate 
each of the likelihoods which will be critical in making our predictions.
\vspace{12 pt}

```{r likelihoods}
##Likelihoods given Rep

#by sex
probMaleGivenRep <- probRepGivenMale*probMale/probRep
probFemaleGivenRep <- probRepGivenFemale*probFemale/probRep

#by race
probWhiteGivenRep <- probRepGivenWhite*probWhite/probRep
probBlackGivenRep <- probRepGivenBlack*probBlack/probRep
probHispanicGivenRep <- probRepGivenHispanic*probHispanic/probRep
probAsianGivenRep <- probRepGivenAsian*probAsian/probRep

#by generation
probMillenialGivenRep <- probRepGivenMillenial*probMillenial/probRep
probGenXGivenRep <- probRepGivenGenX*probGenX/probRep
probBoomerGivenRep <- probRepGivenBoomer*probBoomer/probRep
probSilentGivenRep <- probRepGivenSilent*probSilent/probRep

#by community type
probUrbanGivenRep <- probRepGivenUrban*probUrban/probRep
probSuburbanGivenRep <- probRepGivenSuburban*probSuburban/probRep
probRuralGivenRep <- probRepGivenRural*probRural/probRep

#by income
prob75plusGivenRep <- probRepGiven75plus*prob75plus/probRep
prob30to75GivenRep <- probRepGiven30to75*prob30to75/probRep
problessthan30GivenRep <- probRepGivenlessthan30*problessthan30/probRep

#by education
probPostgradGivenRep <- probRepGivenPostgrad*probPostgrad/probRep
probCollegeGradGivenRep <- probRepGivenCollegeGrad*probCollegeGrad/probRep
probSomeCollegeGivenRep <- probRepGivenSomeCollege*probSomeCollege/probRep
probHighSchoolGivenRep <- probRepGivenHighSchool*probHighSchool/probRep

#by marital status
probMarriedGivenRep <- probRepGivenMarried*probMarried/probRep
probUnmarriedGivenRep <- probRepGivenUnmarried*probUnmarried/probRep

##Likelihoods given Dem

#by sex
probMaleGivenDem <- probDemGivenMale*probMale/probDem
probFemaleGivenDem <- probDemGivenFemale*probFemale/probDem

#by race
probWhiteGivenDem <- probDemGivenWhite*probWhite/probDem
probBlackGivenDem <- probDemGivenBlack*probBlack/probDem
probHispanicGivenDem <- probDemGivenHispanic*probHispanic/probDem
probAsianGivenDem <- probDemGivenAsian*probAsian/probDem

#by generation
probMillenialGivenDem <- probDemGivenMillenial*probMillenial/probDem
probGenXGivenDem <- probDemGivenGenX*probGenX/probDem
probBoomerGivenDem <- probDemGivenBoomer*probBoomer/probDem
probSilentGivenDem <- probDemGivenSilent*probSilent/probDem

#by community type
probUrbanGivenDem <- probDemGivenUrban*probUrban/probDem
probSuburbanGivenDem <- probDemGivenSuburban*probSuburban/probDem
probRuralGivenDem <- probDemGivenRural*probRural/probDem

#by income
prob75plusGivenDem <- probDemGiven75plus*prob75plus/probDem
prob30to75GivenDem <- probDemGiven30to75*prob30to75/probDem
problessthan30GivenDem <- probDemGivenlessthan30*problessthan30/probDem

#by education
probPostgradGivenDem <- probDemGivenPostgrad*probPostgrad/probDem
probCollegeGradGivenDem <- probDemGivenCollegeGrad*probCollegeGrad/probDem
probSomeCollegeGivenDem <- probDemGivenSomeCollege*probSomeCollege/probDem
probHighSchoolGivenDem <- probDemGivenHighSchool*probHighSchool/probDem

#by marital status
probMarriedGivenDem <- probDemGivenMarried*probMarried/probDem
probUnmarriedGivenDem <- probDemGivenUnmarried*probUnmarried/probDem

##Likelihoods given Ind

#by sex
probMaleGivenInd <- probIndGivenMale*probMale/probInd
probFemaleGivenInd <- probIndGivenFemale*probFemale/probInd

#by race
probWhiteGivenInd <- probIndGivenWhite*probWhite/probInd
probBlackGivenInd <- probIndGivenBlack*probBlack/probInd
probHispanicGivenInd <- probIndGivenHispanic*probHispanic/probInd
probAsianGivenInd <- probIndGivenAsian*probAsian/probInd

#by generation
probMillenialGivenInd <- probIndGivenMillenial*probMillenial/probInd
probGenXGivenInd <- probIndGivenGenX*probGenX/probInd
probBoomerGivenInd <- probIndGivenBoomer*probBoomer/probInd
probSilentGivenInd <- probIndGivenSilent*probSilent/probInd

#by community type
probUrbanGivenInd <- probIndGivenUrban*probUrban/probInd
probSuburbanGivenInd <- probIndGivenSuburban*probSuburban/probInd
probRuralGivenInd <- probIndGivenRural*probRural/probInd

#by income
prob75plusGivenDem <- probDemGiven75plus*prob75plus/probDem
prob30to75GivenDem <- probDemGiven30to75*prob30to75/probDem
problessthan30GivenDem <- probDemGivenlessthan30*problessthan30/probDem

#by education
probPostgradGivenInd <- probIndGivenPostgrad*probPostgrad/probInd
probCollegeGradGivenInd <- probIndGivenCollegeGrad*probCollegeGrad/probInd
probSomeCollegeGivenInd <- probIndGivenSomeCollege*probSomeCollege/probInd
probHighSchoolGivenInd <- probIndGivenHighSchool*probHighSchool/probInd

#by marital status
probMarriedGivenInd <- probIndGivenMarried*probMarried/probInd
probUnmarriedGivenInd <- probIndGivenUnmarried*probUnmarried/probInd

```

So, if we come up with an example person, we can see which class (political 
leaning) they are most likely to fall into! Let's suppose that we are interested 
in knowing a black woman's leaning. It's as simple as multiplying the prior for 
a given class by the likelihoods for that person given that class. The class with 
the highest outcome is our prediction! Let's do the calculations.
\vspace{12 pt}

```{r Black_Woman_Prediction}
#Republican

probRep*probBlackGivenRep*probFemaleGivenRep

#Democrat

probDem*probBlackGivenDem*probFemaleGivenDem

#Independent

probInd*probBlackGivenInd*probFemaleGivenInd
```

As we can see, the value for leaning democratic is higher than all of the other
values, so we would predict that a black woman leans democratic based on this 
simple model!
\vspace{12 pt}